--- 
title: "Supplementary materials for manuscript: Differential design intent for Perdiz arrow points from Caddo burial contexts in two landscapes of practice"
author: "Robert Z. Selden, Jr. and John E. Dockall"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Morphology of Perdiz arrow points."
---

# Preface {-}

>Reproducibility---the ability to recompute results---and replicability---the chances other experimenters will achieve a consistent result---are two foundational characteristics of successful scientific research [@RN20918].

## Basis of inquiry

Recent research in the ancestral Caddo area has yielded evidence for a _spatial boundary_ or _behavioural region_, across which material culture from Caddo burials---Hickory Engraved and Smithport Plain bottles as well as Gahagan bifaces---have been found to express significant morphological differences. This inquiry asks whether Perdiz arrow points from Caddo burials, assumed to reflect design intent, may differ across the same _spatial boundary_, extending the pattern of shape differences to a third category of Caddo material culture. Perdiz arrow points collected from the geographies of the northern and southern Caddo communities of practice defined in a social network analysis were employed to test the hypothesis that morphological attributes differ, and are predictable, between the two communities. Results indicate significant between-community differences in maximum length, width, stem length, and stem width, but not thickness. Using the same traditional metrics combined with the tools of machine learning, a predictive model---support vector machine---was designed to assess the degree to which community differences could be predicted, achieving a receiver operator curve score of 97 percent, and an accuracy score of 94 percent. The subsequent geometric morphometric analysis identified significant differences in Perdiz arrow point shape and size, coupled with significant results for modularity and morphological integration, illustrating distinct morphological attributes for two discrete manufacturing traditions expressed primarily in basal morphology between the northern and southern Caddo landscapes of practice.

## Primary findings

### Linear morphometrics

* Maximum length **differs significantly** by Caddo community
* Maximum width **differs significantly** by Caddo community
* Maximum thickness does not differ by Caddo community
* Maximum stem width **differs significantly** by Caddo community
* Maximum stem length **differs significantly** by Caddo community

### Machine learning

* Model achieved a **cross validation score of 86 percent**
* Model achieved a **receiver operator curve score of 97 percent**
* Model achieved an **accuracy score of 94 percent**

### Geometric morphometrics

* Perdiz arrow point shape **differ significantly** by Caddo community
* Perdiz arrow point size **differs significantly** by Caddo community
* Perdiz arrow points exhibit **significant** modularity
* Perdiz arrow points exhibit **significant** blade/base morphological integration

## Acknowledgments

I express my gratitude to the Caddo Nation of Oklahoma and the Anthropology and Archaeology Laboratory at Stephen F. Austin State University for the requisite permissions and access to the NAGPRA items from the Washington Square Mound site and Turner collection, and to Tom A. Middlebrook for brokering access to the Perdiz arrow points from Caddo burials at the Morse Mound site. Thanks also to David K. Thulman, Christian S. Hoggard, Dean C. Adams, Michael L. Collyer, Emma Sherratt, Lauren Butaric, and Kersten Bergstrom for their constructive criticisms, comments, and suggestions throughout the development of this research program, and to the editors and anonymous reviewers for their comments and constructive criticisms, which further improved the manuscript.

## Funding

Components of the analytical workflow were developed and funded by a Preservation Technology and Training grant (P14AP00138) to RZS from the National Center for Preservation Technology and Training, as well as grants to RZS from the Caddo Nation of Oklahoma, National Forests and Grasslands in Texas (15-PA-11081300-033) and the United States Forest Service (20-PA-11081300-074).

## Data management

This volume is written using _Markdown_, and includes all analysis code employed in the study, providing a means for others to reproduce (exactly) those results discussed and expounded upon in the following chapters. The replicable nature of this undertaking provides a means for others to critically assess and evaluate the various analytical components of this study [@RN20915;@RN20916;@RN20917], which is a necessary requirement for the production of reliable knowledge.

Reproducibility projects in [psychology](https://osf.io/ezcuj/) and [cancer biology](https://www.cos.io/rpcb) are impacting current research practices across all domains. Examples of reproducible research are becoming more abundant in archaeology [@RN20804;@RN21009;@RN8322], and the next generation of archaeologists are learning those tools and methods needed to reproduce and/or replicate research results [@RN21007]. Reproducible and replicable research work flows are often employed at the highest levels of humanities-based inquiries to mitigate concern or doubt regarding proper execution, and is of particular import should the results have—explicitly or implicitly—a major impact on scientific progress [@RN21008].

Components of the undertaking presented in this volume are novel, and this basic research endeavour is rooted in archaeological epistemology. Sharing data is not enough, and if investigators share only data absent their analysis code, that places a substantive burden on those who may seek to build upon or replicate their work in the future. In this instance, the choice to pursue a replicable work flow occurs based on the simple fact that this work is foundational, and meant to begin a meaningful discussion related to projectile morphology.

The data and analysis code associated with this project can be accessed through this document or the [GitHub](https://seldenlab.github.io/perdiz3/) repository, and is digitally curated on the Open Science Framework ([DOI 10.17605/OSF.IO/VZHJR](https://osf.io/vzhjr/)).

## Colophon

This version of the analysis was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# what R packages and versions were used?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

Current Git commit details are:

```{r}
# where can I find this commit? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
